---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# HexSampleR

<!-- badges: start -->
<!-- badges: end -->

The **HexSampleR** package is designed to facilitate the creation of sample point locations that 
represent a specific land area such as the area sampled by a single plot within the Forest 
Inventory and Analysis program of the USDA Forest Service.

## Installation

**HexSampleR** is only available as a development version. The **devtools** package is required to
install **HexSampleR**. If you have not previously used **devtools**, use the commented line of
code to install the package.

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
library(devtools)
devtools::install_github("bmcgaughey1/HexSampleR")
```
## Example

This example produces a set of sample points for Snohomish County, Washington. The points represent a 
hexagonal grid with some random offsets. the example uses the *computeCellSize()* helper function to 
compute the cell size needed to generate the sample points given a desired sample area (6000 acres)
and sampling pattern. Note that the units for *cellsize* will be feet since the sample area is in 
acres so we need to explicitly convert to meters to match the UTM projection by dividing by 3.2808.

```{r example, message=FALSE, warning=FALSE}
library(sf)
library(raster)
library(ggplot2)
library(HexSampleR)

# load a polygon representing Snohomish County Washington and reproject from longlat to UTM 10N
WASnoCo <- WASnoCo

# reproject into UTM zone 10...generatePolygonSamplePoints() won't work with data in LAT-LON
WASnoCo <- st_transform(WASnoCo, crs = 26910)

# Creates a set of sample points on a regular hexagonal grid with each point representing
# approximately 6000 acres. This is the area that corresponds to the area sampled by a single FIA plot.
# Offset the points by a random amount to produce a slightly irregular hexagonal grid.
pattern <- "hexagonal"
cellsize <- computeCellSize(6000, aunits = "acre", pattern = pattern) / 3.2808

sample_pts <- generatePolygonSamplePoints(WASnoCo
                                          , polyBuffer = cellsize / 2
                                          , pattern = pattern
                                          , gridSpacing = cellsize
                                          , IDColumnLabel = "NAME"
                                          , perturb = TRUE
                                          , perturbMultiplier = 0.1
                                          )

ggplot(WASnoCo) +
  ggtitle("Hexagonal Grid of Sample Points With Random Offset") +
  geom_sf(aes(fill = COUNTYFP), show.legend = FALSE) +
  geom_sf(data = sample_pts, colour = "black", size = 2)
```
